Selective stock enhancement undermines community stability

### 2022-01-19 rmd/manuscript.Rmd
edit: manuscript

# theory infomation
df_theory <- read_csv(here::here("code_theory/result/result_ricker.csv"))

@@

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to the sustainable management of natural resources [@cardinaleBiodiversityLossIts2012]. Captive breeding is a common practice to conserve or enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996]. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles) [@arakiGeneticEffectsCaptive2007a; @snyderLimitationsCaptiveBreeding1996; @griffithsCaptiveBreedingReintroduction2008], this method is still pervasive in conservation and fisheries. In fisheries for example, billions of hatchery-reared individuals are produced and released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species through density-dependent feedback underpins emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999a]. Stock enhancement may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild. Hence, this form of species management may have significant consequences for community dynamics. Evidence for this hypothesis is lacking, however.

Here, we use a simple theory to show that stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999a]. Our theory illuminates that stock enhancement interferes with the internal regulation process of self-sustainable communities, compromising the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by illustrating its congruence with Japanese stream fish communities, where \>10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

@@

$$
N_{i,t+1} = (N_{i,t}+\phi_i R_t)~exp[r_i(1-\frac{N_{i,t}+\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K})]~exp(\epsilon_{i,t})
$$

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ (interspecific competition drawn from an exponential distribution as $\alpha_{ij,i \ne j} \sim Exp(1/\bar{\alpha})$ whereas intraspecific competition $\alpha_{ii}=1$), $K$ the community-level carrying capacity in the absence of niche difference, $R_t$ the number of released individuals for the enhancement of species 1 (set constant as $R_t = R$ for simplicity), and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations that follow an normal distribution with a mean of zero and a standard deviation of $\sigma_{\epsilon}$. The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

@@

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Therefore, the model accounts for fitness difference of captive-bred individuals due to genetic effects and/or phenotypic plasticity when considering the reproductive contribution of released individuals to the next generation of the enhanced species.

In our simulation, we used `r nrow(param_set)` simulation scenarios with varied values of the following ecological parameters: mean strength of interspecific competition $\bar{\alpha}$, intrinsic growth of the enhanced species $r_1$, relative fitness of captive-bred individuals $f_R$, and degree of environmental stochasticity $\sigma_{\epsilon}$. Under each scenario with a unique parameter combination, we ran `r n_sim` time steps of `r max(df_theory$n_rep)` independent communities (i.e., simulation replicates) with varying numbers of released individuals $R$. The last `r unique(df_theory$n_timestep)` time steps were used to obtain the coefficient of variation, temporal mean, and temporal standard deviation of total community density ($\sum_i^SN_i$). Temporal means and standard deviations were calculated separately for the enhanced species $N_1$ and the unenhanced species $\sum_{i,i \ne 1}^S N_i$ to infer mechanisms that underlie community-level responses.

Our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced means of enhanced and unenhanced species groups (Figure XX) that are driven by different mechanisms. For enhanced species, released individuals induced overcompensation, where the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals [@satakeStockingCaptivebredFish2012]. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals. Meanwhile, interspecific competition between enhanced and unenhanced species caused the reduced mean of unenhanced species, leading to decreased numbers of persisting species at high levels of stock enhancement. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density has led to the increased CV. These patterns were qualitatively similar across ecological scenarios, even when the interspecific competition is very weak.

The destabilizing effect emerges because stock enhancement undermines the balance of species interactions that underpin community stability. Stable coexistence of competing species rests largely on niche difference in consumable resource use ($\alpha_{ij, i \ne j} < 1$) because it creates a situation where numerically dominant species undergoes stronger intraspecific competition [@lettenLinkingModernCoexistence2017]. However, stock enhancement is externally controlled and the number of release is not subject to density-dependent regulation. Such excessive individuals impose additional intra- and interspecific competition. Consequently, stock enhancement compromised the ecological process giving rise to "overyielding" [@tilmanEcologicalConsequencesChanges1999a], in which total community density of an ecologically diverse community exceeds what would be expected in a single species community ($\sum_i^SN_i > K$).

We must note that stock enhancement helps increase the density of the enhanced species when the intrinsic population growth rate is significantly small ($r_1 = 0.5$). This result is reasonable because some of the best evidence for successful stock enhancement comes from critically endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. Yet, the community-level effect is still destabilizing due to the intensified competition with unenhanced species. It is therefore critical to consider the propagating effect when implementing captive breeding programs.

@@

In our theory, the observed pattern emerges when negative density dependence operates in the recipient community, and this assumption is well supported by previous observations. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that released masu salmon compete with wild individuals of masu salmon [@sahashiEvaluationContributionHatchery2015a] and other stream fishes [@hasegawaEffectsHatcheryChum2014; @hasegawaWildMasuSalmon2018]. In particular, masu salmon raised in hatchery environments are larger and more aggressive than wild individuals, increasing the likelihood of intense intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018]. Importantly, the number of fish released is likely controlled by socioeconomic factors (e.g., budget allocation) because there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). As such, released fish are "excessive" given the density-dependent regulation of recipient communities and may impose resource competition that would otherwise not exist.

Reduced fitness of captive-bred individuals may also play a role in producing the observed community response. As with other salmonid species, ocean-migrating adults of captive-bred masu salmon show lower return rates to the spawning river and therefore less contribute to the next generation (miyakoshi?). These processes, i.e., negative density dependence and fitness disadvantage of captive-bred fish, may act in concert to reduce the long-term average of community density, undermining community stability in the long term.

@@

Although captive breeding programs pay close attention to the fate of captive-bred individuals in the wild, it is rare to consider the self-regulation process of biodiversity. The results of this study demonstrate that the ignorance of this critical process will erode the long-term persistence of the recipient community. The destabilizing effect of stock enhancement is likely to be pervasive across terrestrial and aquatic taxa because negative density dependence is a general regulating mechanism of ecological communities.

We anticipate that the undesired influence of stock enhancement may be more apparent in degraded habitats as the compromised capacity to support biodiversity may impose greater competition among constituent species. Therefore, we should shift our management efforts to the preservation and restoration of natural habitats with a broader scope to conserve biodiversity. For example, a modern conservation initiative underscores the importance of conserving a diverse habitat portfolio that may buffer the impacts of rapid environmental changes. Without a comprehensive framework that appreciates the ecological integrity of self-sustaining communities, stock enhancement will never be effective but impairs biodiversity.

>>>>

# theory infomation
df_theory <- read_csv(here::here("result/result_ricker.csv"))

@@

Anthropogenic influences have caused global biodiversity crisis, such that human interventions are critical to the sustainable management of natural resources [@cardinaleBiodiversityLossIts2012]. Captive breeding is a common practice to conserve or enhance wild populations of diverse plant and animal taxa [@snyderLimitationsCaptiveBreeding1996]. Although there is growing recognition that stock enhancement with captive-bred organisms entails ecological risks (e.g., the accumulation of deleterious alleles) [@arakiGeneticEffectsCaptive2007a; @krkosekDecliningWildSalmon2007], this method is still pervasive in conservation [@snyderLimitationsCaptiveBreeding1996; @griffithsCaptiveBreedingReintroduction2008; @haagNorthAmericanFreshwater2012] and fisheries [@kitadaEconomicEcologicalGenetic2018; @arakiHatcheryStockingHelp2010]. In fisheries for example, billions of hatchery-reared individuals are released annually across the globe [@kitadaEconomicEcologicalGenetic2018]. The widespread use of captive breeding is perhaps because of the "myth" that numerical benefits of released individuals may exceed the associated risks.

Current debates, however, overlook the fact that we have rarely assessed the community-wide impact of stock enhancement. Species are all embedded in the complex web of interacting organisms, and stable coexistence of competing species through density-dependent feedback underpins emergent stability of ecological communities [@tilmanEcologicalConsequencesChanges1999a; @carrollNicheFitnessDifferences2011a; @loreauDoesFunctionalRedundancy2004]. Stock enhancement may disrupt the sensitive balance of species interactions because it introduces unnaturally high numbers of individuals into the wild [@weberInteractionsHatcheryWild2003; @careyNativeInvadersChallenges2012]. Hence, this form of species management may have significant consequences for community dynamics in the long term. Evidence for this hypothesis is lacking, however.

Here, we use a simple theory to show that stock enhancement undermines long-term community stability, which we define here as the relative size of fluctuations in total community density over time [@tilmanEcologicalConsequencesChanges1999a]. Our theory illuminates that stock enhancement interferes with the internal regulation process of self-sustainable communities, compromising the stabilizing mechanism emerging from species niche differences. The present study further demonstrates the relevance of this general theory to natural systems by showing its congruence with Japanese stream fish communities, where \>10 million hatchery masu salmon (*Oncorhynchus masou masou*) are released annually for fisheries, recreation, and conservation purposes [@kitadaLessonsJapanMarine2020]. Our findings paint a bleak picture for stock enhancement, reinforcing the recurrent calls for alternative management strategies.

@@

$$
N_{i,t+1} = (N_{i,t}+\phi_i R_t)~exp[r_i(1-\frac{N_{i,t}+\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i})]~exp(\epsilon_{i,t})
$$

where $r_i$ is the intrinsic growth rate (drawn from a uniform distribution as $r_i \sim Unif(0,2)$ except $r_1$), $\alpha_{ij}$ the competition coefficient of species $j$ on species $i$ [drawn from an exponential distribution as $\alpha_{ij,i \ne j} \sim Exp(1/\bar{\alpha})$; intraspecific competition $\alpha_{ii}=1$], $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations that follow an normal distribution with a mean of zero and a standard deviation of $\sigma_{\epsilon}$. The parameter $\phi_i$ controls the relative fitness of captive-bred individuals as follows:

@@

$f_R~(\ge 0)$ is the density-independent survival of captive-bred individuals relative to wild individuals. Therefore, the model accounts for fitness difference of captive-bred individuals due to genetic effects and/or plasticity [@weberInteractionsHatcheryWild2003; @snyderLimitationsCaptiveBreeding1996; @arakiHatcheryStockingHelp2010] when considering the reproductive contribution of released individuals to the next generation of the enhanced species. Without loss of generality, we assumed constant $K_i$ ($K_i = K$) and $R_t$ ($R_t = R$) across species and time, respectively.

In our simulation, we used `r nrow(param_set)` simulation scenarios with varied values of the following ecological parameters: intrinsic growth of the enhanced species $r_1$, mean strength of interspecific competition $\bar{\alpha}$, carrying capacity $K$, relative fitness of captive-bred individuals $f_R$, and degree of environmental stochasticity $\sigma_{\epsilon}$. Under each scenario with a unique parameter combination, we ran `r n_sim` time steps of `r max(df_theory$n_rep)` independent communities (i.e., simulation replicates) with varying numbers of released individuals $R$ ($0-500$). The last `r unique(df_theory$n_timestep)` time steps were used to obtain the coefficient of variation, temporal mean, and temporal standard deviation of total community density ($\sum_i^SN_i$). Temporal means and standard deviations were calculated separately for the enhanced species $N_1$ and unenhanced species $\sum_{i,i \ne 1}^S N_i$ to infer mechanisms that underlie the community-level response.

In general, our model predicted a destabilizing effect of selective stock enhancement on ecological communities as illustrated by increased CV with increasing numbers of released individuals $R$ (Figure XX). This pattern stemmed mainly from the reduced mean of total community density, and both enhanced and unenhanced species groups were responsible for this result. For enhanced species, released individuals induced overcompensation, in which the negative effect of reduced per-capita reproduction due to competition exceeds the numerical benefit of released individuals. Therefore, the enhanced species reached the lower long-term average of population density with higher numbers of released individuals, confirming previous theoretical predictions [@satakeStockingCaptivebredFish2012]. Meanwhile, interspecific competition caused the reduced mean density of unenhanced species and led to decreased numbers of persisting species at high levels of stock enhancement. Combined, the long-term average of the total community density decreased more sharply than individual species groups with stock enhancement (XX). The SDs showed a similar trend, but the relationship was flatter at the community level. Since a CV is a ratio of an SD to a mean, the steeper relationship of the mean community density resulted in the increased CV. These patterns were qualitatively similar across most ecological scenarios, even when the interspecific competition is very weak (XX).

The destabilizing effect emerges because stock enhancement undermines the balance of species interactions that underpin community stability. Ecological theory suggests that stable coexistence of competing species requires a niche difference that is large enough to overcome the relative difference in intrinsic competitive ability [@carrollNicheFitnessDifferences2011a]. Under this condition, competing species can grow from small populations because numerically dominant species undergoes stronger intraspecific competition [@carrollNicheFitnessDifferences2011a]. Such stable coexistence increases the likelihood of stabilizing temporal dynamics of species rich communities [@tilmanEcologicalConsequencesChanges1999a] because it gives rise to "overyielding" [@loreauDoesFunctionalRedundancy2004; @carrollNicheFitnessDifferences2011a], that is, total community density of a multi-species community exceeds what would be expected in a single species community ($\sum_i^S N_i > K$). However, stock enhancement is externally controlled and the number of release is not subject to density-dependent regulation. Therefore, excessive individuals supplemented through stock enhancement impose additional intra- and interspecific competition that interferes with the ecological process producing overyielding.

The only scenario where stock enhancement had little influences on community dynamics was when carrying capacity was sufficiently large ($K=500$). In particular, stock enhancement increased, albeit slightly, the average density of the enhanced species with low population growth rate ($r_1=0.5$). This result is reasonable because some of the best evidence for successful stock enhancement comes from critically endangered species [@johnsonGeneticRestorationFlorida2010; @oppelPopulationReinforcementDemographic2021]. In the meantime, our theoretical analysis also illuminates limitation of stock enhancement as a primary tool of ecosystem management. Habitat loss and fragmentation have severely reduced biodiversity, and one of the primary motivations for stock enhancement is to mitigate or reverse such a trend. However, stock enhancement will not bring desired ecological outcomes unless we address the root causes of habitat degradation and restore environmental capacity to support biodiversity.

@@

In our theory, the observed pattern emerges when negative density dependence operates in the recipient community, and this assumption is well supported by previous observations. Masu salmon typically spend one year in freshwater before migrating to the ocean, and a wealth of evidence suggests that released masu salmon compete with wild individuals of masu salmon [@sahashiEvaluationContributionHatchery2015a] and other stream fishes [@hasegawaEffectsHatcheryChum2014; @hasegawaWildMasuSalmon2018]. Hatchery-reared salmons are larger and more aggressive than wild individuals, further increasing the likelihood of intense intra- and interspecific competition in the wild [@hasegawaWildMasuSalmon2018; @weberInteractionsHatcheryWild2003]. Importantly, the number of fish released is likely controlled by socioeconomic factors (e.g., budget allocation) because there was no apparent congruence between the number of fish released and the population density of masu salmon (XX). As such, released fish are "excessive" given the density-dependent regulation of recipient communities and may impose resource competition that would otherwise not exist.

Reduced fitness of captive-bred individuals may also play a role in producing the observed community response. As with other salmonid species, ocean-migrating adults of captive-bred masu salmon show lower return rates to the spawning river and therefore less contribute to the next generation (miyakoshi?). It has also been shown that hatchery-reared salmons are more susceptible to predation than wild salmons [@weberInteractionsHatcheryWild2003]. The fitness disadvantage of captive-bred fish may interact with negative density dependence to reduce the long-term average of community density, undermining community stability in the long term.

@@

The past few decades have witnessed substantial efforts to unveil ecological consequences of captive breeding, and there is a general agreement that the distinctive features of captive-bred individuals bring fitness disadvantages in the wild [@arakiHatcheryStockingHelp2010; @snyderLimitationsCaptiveBreeding1996]. As such, numerous studies suggested that stock enhancement with captive breeding must be carefully employed given the potential negative impacts on wild populations of enhanced species [@christieGeneticAdaptationCaptivity2012; @arakiGeneticEffectsCaptive2007a]. Despite the significant attention to the fate of captive-bred individuals, it is rare to consider the self-regulation process of biodiversity. The results of this study suggest that the ignorance of this critical process may erode the long-term persistence of the recipient community. While our empirical example is limited to stream fish communities, the consistent emergence of destabilizing effect across simulated ecological scenarios implies that this ecological phenomenon is pervasive across terrestrial and aquatic taxa.

We should prioritize the preservation and restoration efforts of natural habitats with a broader scope of biodiversity conservation. In particular, such efforts are crucial given that the undesired influence of stock enhancement is predicted to be exacerbated in degraded habitats due to the compromised capacity to support biodiversity (XX). It has been shown that protected areas and environmental restoration are promising tools to conserve biodiversity, and the effective spatial design is a key to achieve successful conservation [@koningNetworkGrassrootsReserves2020; @akasakaSmartProtectedArea2017]. For example, coordinated placement of conservation sites considering spatial biodiversity patterns is of critical importance in improving the ecological outcomes [@akasakaSmartProtectedArea2017; @teruiEmergentDualScaling2021; @koningNetworkGrassrootsReserves2020; @stollSmallImpoverishedRegional2013]. Socioeconomic factors such as governance may also play a central role in enforcing environmental legislation, potentially determining the effectiveness of conservation investment [@amanoSuccessfulConservationGlobal2018]. These considerable potentials indicate that we have viable management options for the sustainable future of biodiversity and natural resources before blindly accepting stock enhancement programs. Without a comprehensive framework that appreciates the ecological integrity of natural communities, stock enhancement will never be effective but impairs biodiversity.
